---
title: "Written Exams - DeMarco - Aug 2022"
output:
  html_document:
    df_print: paged
---
Data: Provided from ICPSR_36873-V5 as '36873-0001-Data.rda'
Codebook: 38673-0001-Codebook.pdf
Homepage of study site: https://www.icpsr.umich.edu/web/ICPSR/studies/36873

Research Question:
Among older adult respondents to the NSHAP, do frequent anxiety and depression symptoms impact the severity of functional limitations in activities of daily living and instrumental activities of daily living?

Packages required include:
Tidyverse, tableone, survey, ggsurvey, extrafont, prettyR, svyVGAM
```{r Load packages}
# Create a list of required packages
packages <- c("tidyverse", "tableone", "survey", "ggsurvey", "extrafont", "prettyR", "svyVGAM", "jtools")

# Load the required packages
purrr::walk(packages, library, character.only = T)

# Display citations for the required packages
for (package in packages){
  print(citation(package = package))}
```

```{r Load data}
load(file = "C:\\Users\\samzo\\Downloads\\ICPSR_34921-V5\\ICPSR_34921\\DS0001\\34921-0001-Data.rda")
```

The "factor_to_numeric_icpsr.R" script file provides the necessary code to turn the factor variables into a numeric variable. This is turned into a function for use in the present analysis.
```{r Factor to Numeric}
fac_to_numeric <- function(dfvar){
  lbls <- sort(levels(dfvar))
  lbls <- (sub("^\\([0-9]+\\) +(.+$)", "\\1", lbls))
  newvar <- as.numeric(sub("^\\(0*([0-9]+)\\).+$", "\\1", dfvar))
  newvar <- add.value.labels(newvar, lbls)
  return(newvar)
}
```

Testing the new function to see if it works with tidyverse
```{r testing new function}
testdf <- da34921.0001 %>%
  mutate(PrepMealRecode = fac_to_numeric(PREPMEAL))%>%
  select(PrepMealRecode)
summary(testdf)
```

The study sample will be limited to those who responded to both the frequent anxiety symptoms and frequent depression symptoms measures, had documented responses for both the ADL and IADL measures, and could do each ADL/IADL at some point (a.k.a. did not volunteer "have never done" for any activity).

The exposure of interest is a composite measure of mental health with 4 levels - those with neither frequent anxiety symptoms nor frequent depressive symptoms, those with frequent anxiety symptoms in the absence of frequent depressive symptoms, those with frequent depressive symptoms in the absence of frequent anxiety symptoms, and those with both frequent anxiety and depressive symptoms. 

The primary outcome of interest is functional ability as measured by the severity of difficulty with activities of daily living and instrumental activities of daily living (Cheung, 2021). Severity was assessed at 4 levels - no difficulty (0), some difficulty (1), much difficulty (2), and unable to do (3). A higher score indicates greater difficulty with ADLs and IADLs.

Secondary outcomes include the count of ADLs and IADLs with which a respondent requires assistance and a composite variable of the type of assistance needed (none, IADL only, ADL only, both IADL and ADL). Previously published scales for the NSHAP ADL and IADL data (Huisingh-Scheetz, 2014) will be used. [NOTE: this is actually backwards from the scale used and instead is better described as assessing/finding individuals with any limitation] A point is awarded for any level of difficulty with each of 7 key activities of daily living (walking a block, walking across the room, dressing, bathing, eating, transferring in and out of bed, and toileting) and 8 instrumental activities of daily living (preparing meals, taking medication, managing money, shopping for food, performing light housework, using a telephone, driving during the day, and driving at night). 

Covariates include self-perceived physical health, number of comorbidities (count of those assessed by NSHAP), presence of health insurance, age gender, income, marital status, and education.

An analytic data set is created with the necessary variables.
```{r Clean Data}


analyticNSHAP <- da34921.0001 %>%
  # Change depressive symptoms to numeric for ease of analysis
  mutate(noteat = fac_to_numeric(NOTEAT))%>%
  mutate(feltdep = fac_to_numeric(FLTDEP))%>%
  mutate(felteff = fac_to_numeric(FLTEFF))%>%
  mutate(nosleep = fac_to_numeric(NOSLEEP))%>%
  mutate(washappy = fac_to_numeric(WASHAPY))%>%
  mutate(waslonely = fac_to_numeric(WASLONLY))%>%
  mutate(unfriend = fac_to_numeric(UNFRIEND))%>%
  mutate(enjlife = fac_to_numeric(ENJLIFE))%>%
  mutate(feltsad = fac_to_numeric(FLTSAD))%>%
  mutate(disliked = fac_to_numeric(DISLIKD))%>%
  mutate(notgetgo = fac_to_numeric(NOTGETGO))%>%
  # Change anxiety symptoms to numeric for ease of analysis
  mutate(felttens = fac_to_numeric(FLTTENS))%>%
  mutate(fright = fac_to_numeric(FRIGHT))%>%
  mutate(worry = fac_to_numeric(WORRY))%>%
  mutate(relaxed = fac_to_numeric(RELAXED))%>%
  mutate(butterfly = fac_to_numeric(BUTRFLY))%>%
  mutate(restless = fac_to_numeric(RESTLES))%>%
  mutate(panic = fac_to_numeric(PANIC))%>%
  # Change comorbidities to numeric for ease of analysis
  mutate(htn = fac_to_numeric(CONDITNS_6))%>%
  mutate(arthritis = fac_to_numeric(ARTHRITIS))%>%
  mutate(heartcond = fac_to_numeric(HRTPROB2))%>%
  mutate(cancer = fac_to_numeric(OTHCAN))%>%
  mutate(diabetes = fac_to_numeric(CONDITNS_7))%>%
  mutate(lung = fac_to_numeric(EMPHASTH))%>%
  mutate(stroke = fac_to_numeric(STROKE))%>%
  mutate(cognition = fac_to_numeric(DEMENTIA))%>%
  # Create a summary variable for number of comorbidities
  mutate(comorbid = htn+arthritis+heartcond+ cancer+ diabetes+lung+stroke+cognition)%>%
  # Select necessary variables
  select(ID, WEIGHT_ADJ, STRATUM, CLUSTER, AGE, GENDER, ETHGRP, EDUC, MARITLST, PHYSHLTH, comorbid,
         MNTLHLTH, HAPPY, HEARN_RECODE, OTHERINSUR, 
         WALKBLK, WALKBLOCK_HELP, WALKROOM, WALKROOM_HELP, DRESSING, DRESS_HELP, BATHING, BATH_HELP,
         EATING, EATING_HELP, INOUTBED, GETBED_HELP, TOILET, TOILET_HELP, PREPMEAL, TAKEMEDS, MANAGEMONEY, SHOPFOOD, LIGHTWORK, USETELEP, DRIVED2, DRIVEN2, noteat, feltdep, felteff,
         nosleep, washappy, waslonely, unfriend, enjlife, feltsad, disliked, notgetgo, felttens, 
         fright, worry, relaxed, butterfly, restless, panic)
```


Primary outcome is severity of functional limitation as measured by difficulty with ADLs and IADLs (Cheung, 2021). Those who reporting having never done an activity were re-coded as not applicable as their baseline function is likely different from the majority of participants. A new function was created to apply these rules to all ADL and IADL variables, as the coding scheme was the same.
```{r ADL recode function}
ADLrecode <- function(adlvar){
  newvar <- case_when(adlvar == "(0) no difficulty" ~ 0,
                            adlvar == "(4) if volunteered - have never done" ~ NA,
                            adlvar == "(1) some difficulty" ~ 1,
                            adlvar == "(2) much difficulty" ~ 2,
                            adlvar == "	
(3) unable to do" ~ 3)
  return(newvar)
}
```

The severity of ADL limitation was assessed by constructing a scale as described above using the 7 ADLs assessed.
```{r severity of ADL need}

analyticNSHAP <- analyticNSHAP %>%
  # Create severity variables by using the ADLrecode function
  mutate(block_severity = ADLrecode(WALKBLK))%>%
  mutate(room_severity = ADLrecode(WALKROOM))%>%
  mutate(dress_severity = ADLrecode(DRESSING))%>%
  mutate(bathe_severity = ADLrecode(BATHING))%>%
  mutate(eat_severity = ADLrecode(EATING))%>%
  mutate(bed_severity = ADLrecode(INOUTBED))%>%
  mutate(toilet_severity = ADLrecode(TOILET))%>%
  # Calculate an overall severity score by summing severity of difficulty for tasks
  mutate(ADL_severity = block_severity + room_severity + dress_severity + bathe_severity + eat_severity + bed_severity + toilet_severity)

# View a summary
summary(analyticNSHAP$ADL_severity)

```

The severity of IADL limitation was assessed by constructing a scale as described above using the 8 IADLs assessed.
```{r severity of IADL need}
analyticNSHAP <- analyticNSHAP%>%
  # create severity variables by using the ADLrecode function
  mutate(meal_severity = ADLrecode(PREPMEAL))%>%
  mutate(meds_severity = ADLrecode(TAKEMEDS))%>%
  mutate(money_severity = ADLrecode(MANAGEMONEY))%>%
  mutate(shop_severity = ADLrecode(SHOPFOOD))%>%
  mutate(housework_severity = ADLrecode(LIGHTWORK))%>%
  mutate(phone_severity = ADLrecode(USETELEP))%>%
  mutate(daydrive_severity = ADLrecode(DRIVED2))%>%
  mutate(nightdrive_severity = ADLrecode(DRIVEN2))%>%
  # calculate an overall severity score by summing severity of difficulty for tasks
  mutate(IADL_severity = meal_severity + meds_severity + money_severity + shop_severity + housework_severity + phone_severity + daydrive_severity + nightdrive_severity)

# view a summary
summary(analyticNSHAP$IADL_severity)
```


The secondary outcome of interest is the type of functional ability as measured by activities of daily living and instrumental activities of daily living. 
```{r Variable definition - need ADL assistance}
analyticNSHAP <- analyticNSHAP %>%
  # Create dichotomous variables for each ADL to determine need for assistance
  mutate(block_binary = case_when(WALKBLK == "(0) no difficulty" ~ 0,
                                  WALKBLK == "(4) if volunteered - have never done" ~ NaN,
                                  WALKBLK == "(1) some difficulty" |  WALKBLK == "(2) much difficulty" |  WALKBLK == "(3) unable to do" ~ 1))%>%
  mutate(room_binary = case_when(WALKROOM == "(0) no difficulty" ~ 0,
                                  WALKROOM == "(4) if volunteered - have never done" ~ NaN,
                                  WALKROOM == "(1) some difficulty" |  WALKROOM == "(2) much difficulty" |  WALKROOM == "(3) unable to do" ~ 1))%>%
  mutate(dress_binary = case_when(DRESSING == "(0) no difficulty" ~ 0,
                                  DRESSING == "(4) if volunteered - have never done" ~ NaN,
                                  DRESSING == "(1) some difficulty" |  DRESSING == "(2) much difficulty" |  DRESSING == "(3) unable to do" ~ 1))%>%
  mutate(bathe_binary = case_when(BATHING == "(0) no difficulty" ~ 0,
                                  BATHING == "(4) if volunteered - have never done" ~ NaN,
                                  BATHING == "(1) some difficulty" |  BATHING == "(2) much difficulty" |  BATHING == "(3) unable to do" ~ 1))%>%
  mutate(eat_binary = case_when(EATING == "(0) no difficulty" ~ 0,
                                  EATING == "(4) if volunteered - have never done" ~ NaN,
                                  EATING == "(1) some difficulty" |  EATING == "(2) much difficulty" |  EATING == "(3) unable to do" ~ 1))%>%
  mutate(bed_binary = case_when(INOUTBED == "(0) no difficulty" ~ 0,
                                  INOUTBED == "(4) if volunteered - have never done" ~ NaN,
                                  INOUTBED == "(1) some difficulty" |  INOUTBED == "(2) much difficulty" |  INOUTBED == "(3) unable to do" ~ 1))%>%
  mutate(toilet_binary = case_when(TOILET == "(0) no difficulty" ~ 0,
                                  TOILET == "(4) if volunteered - have never done" ~ NaN,
                                  TOILET == "(1) some difficulty" |  TOILET == "(2) much difficulty" |  TOILET == "(3) unable to do" ~ 1))%>%
  # Calculate an ADL score based on previous guidelines, with 1 point for any difficulty
  mutate(ADL_score = block_binary + room_binary + dress_binary + bathe_binary + eat_binary + bed_binary + toilet_binary)%>%
  # Variable for needing any ADL assistance, defined as an ADL score >=1
  mutate(needADLhelp = case_when(ADL_score >= 1 ~ "Yes",
                                 ADL_score < 1 ~ "No"))%>%
  mutate(needADLhelp = factor(needADLhelp, levels = c("No", "Yes"), ordered = T))

# View a summary of the new variables
summary(analyticNSHAP$ADL_score)
summary(analyticNSHAP$needADLhelp)
```

Instrumental activities of daily living scale
```{r variable definition - IADL scale}
analyticNSHAP <- analyticNSHAP %>%
  # Create dichotomous variables for each IADL to determine need for assistance
  mutate(meal_binary = case_when(PREPMEAL == "(0) no difficulty" ~ 0,
                                  PREPMEAL == "(4) if volunteered - have never done" ~ NaN,
                                  PREPMEAL == "(1) some difficulty" |  PREPMEAL == "(2) much difficulty" |  PREPMEAL == "(3) unable to do" ~ 1))%>%
  mutate(meds_binary  = case_when(TAKEMEDS == "(0) no difficulty" ~ 0,
                                  TAKEMEDS == "(4) if volunteered - have never done" ~ NaN,
                                  TAKEMEDS == "(1) some difficulty" |  TAKEMEDS == "(2) much difficulty" |  TAKEMEDS == "(3) unable to do" ~ 1))%>%
  mutate(money_binary = case_when(MANAGEMONEY == "(0) no difficulty" ~ 0,
                                  MANAGEMONEY == "(4) if volunteered - have never done" ~ NaN,
                                  MANAGEMONEY == "(1) some difficulty" |  MANAGEMONEY == "(2) much difficulty" |  MANAGEMONEY == "(3) unable to do" ~ 1))%>%
  mutate(shopfood_binary  = case_when(SHOPFOOD == "(0) no difficulty" ~ 0,
                                  SHOPFOOD == "(4) if volunteered - have never done" ~ NaN,
                                  SHOPFOOD == "(1) some difficulty" |  SHOPFOOD == "(2) much difficulty" |  SHOPFOOD == "(3) unable to do" ~ 1))%>%
  mutate(housework_binary  = case_when(LIGHTWORK == "(0) no difficulty" ~ 0,
                                  LIGHTWORK == "(4) if volunteered - have never done" ~ NaN,
                                  LIGHTWORK == "(1) some difficulty" |  LIGHTWORK == "(2) much difficulty" |  LIGHTWORK == "(3) unable to do" ~ 1))%>%
  mutate(phone_binary  = case_when(USETELEP == "(0) no difficulty" ~ 0,
                                  USETELEP == "(4) if volunteered - have never done" ~ NaN,
                                  USETELEP == "(1) some difficulty" |  USETELEP == "(2) much difficulty" |  USETELEP == "(3) unable to do" ~ 1))%>%
  mutate(daydrive_binary = case_when(DRIVED2 == "(0) no difficulty" ~ 0,
                                  DRIVED2 == "(4) if volunteered - have never done" ~ NaN,
                                  DRIVED2 == "(1) some difficulty" |  DRIVED2 == "(2) much difficulty" |  DRIVED2 == "(3) unable to do" ~ 1))%>%
  mutate(nightdrive_binary  = case_when(DRIVEN2 == "(0) no difficulty" ~ 0,
                                  DRIVEN2 == "(4) if volunteered - have never done" ~ NaN,
                                  DRIVEN2 == "(1) some difficulty" |  DRIVEN2 == "(2) much difficulty" |  DRIVEN2 == "(3) unable to do" ~ 1))%>%
  # Create a scored variable as the sum of the binary variables
  mutate(IADL_score = meal_binary + meds_binary + money_binary + shopfood_binary + housework_binary + phone_binary + daydrive_binary + nightdrive_binary) %>%
  # Create a binary variable for needing any IADL help overall
  mutate(IADL_needHelp = case_when(IADL_score >= 1 ~ "Yes",
                                  IADL_score <1 ~ "No"))%>%
  mutate(IADL_needHelp = factor(IADL_needHelp, levels = c("No", "Yes"), ordered = T))

# View a summary of the new variables
summary(analyticNSHAP$IADL_score)
summary(analyticNSHAP$IADL_needHelp)
```

Functional ability can be broadly defined by both ADLs and IADLs. A composite score is created by summing the ADL and IADL score.
```{r Composite ADL IADL Score}
analyticNSHAP <- analyticNSHAP %>%
  # Sum the ADL and IADL scores
  mutate(compADL_score = ADL_score + IADL_score)%>%
  # Create composite ordinal variable
  mutate(compADL_ordinal = case_when(needADLhelp == "No" & IADL_needHelp == "No" ~ "Independent",
                                     needADLhelp == "No" & IADL_needHelp == "Yes" ~ "IADL Help Only",
                                     needADLhelp == "Yes" & IADL_needHelp == "No" ~ "ADL Help Only",
                                     needADLhelp == "Yes" & IADL_needHelp == "Yes" ~ "Both ADL and IADL Help"))%>%
  mutate(compADL_ordinal = factor(compADL_ordinal, levels = c("Independent", "IADL Help Only", "ADL Help Only", "Both ADL and IADL Help"), ordered = T))

# View summary
summary(analyticNSHAP$compADL_score)
summary(analyticNSHAP$compADL_ordinal)
```

ADL difficulty and IADL difficulty can be assessed using a linear scale, where the reported difficulty for each activity wihtin the domain is summed. 


Frequent depressive symptoms is designated as score >= 9. Scoring was conducted following the provided guidelines. (Payne, 2014)
```{r Scoring for frequent depressive symptoms}
analyticNSHAP <- analyticNSHAP %>%
  # Assign scores to each item
  mutate(noteat = case_when(noteat == 1 ~ 0,
                            noteat == 2 ~ 1, 
                            noteat == 3 | noteat == 4 ~ 2))%>%
  mutate(feltdep = case_when(feltdep == 1 ~ 0,
                            feltdep == 2 ~ 1, 
                            feltdep == 3 | feltdep == 4 ~ 2))%>%
  mutate(felteff = case_when(felteff == 1 ~ 0,
                            felteff == 2 ~ 1, 
                            felteff == 3 | felteff == 4 ~ 2))%>%
  mutate(nosleep = case_when(nosleep == 1 ~ 0,
                            nosleep == 2 ~ 1, 
                            nosleep == 3 | nosleep == 4 ~ 2))%>%
  mutate(washappy = case_when(washappy == 1 ~ 2, 
                              washappy == 2 ~ 1,
                              washappy == 3 | washappy == 4 ~ 0))%>%
  mutate(waslonely = case_when(waslonely == 1 ~ 0,
                            waslonely == 2 ~ 1, 
                            waslonely == 3 | waslonely == 4 ~ 2))%>%
  mutate(unfriend = case_when(unfriend == 1 ~ 0,
                            unfriend == 2 ~ 1, 
                            unfriend == 3 | unfriend == 4 ~ 2))%>%
  mutate(enjlife = case_when(enjlife == 1 ~ 0,
                            enjlife == 2 ~ 1, 
                            enjlife == 3 | enjlife == 4 ~ 2))%>%
  mutate(feltsad = case_when(feltsad == 1 ~ 0,
                            feltsad == 2 ~ 1, 
                            feltsad == 3 | feltsad == 4 ~ 2))%>%
  mutate(disliked = case_when(disliked == 1 ~ 0,
                            disliked == 2 ~ 1, 
                            disliked == 3 | disliked == 4 ~ 2))%>%
  mutate(notgetgo = case_when(notgetgo == 1 ~ 0,
                            notgetgo == 2 ~ 1, 
                            notgetgo == 3 | notgetgo == 4 ~ 2))%>%
# Create a total score variable based on the sum
  mutate(fds_score = noteat + feltdep + felteff + nosleep + washappy + waslonely + unfriend + enjlife + feltsad + disliked + notgetgo)%>%
  # Create a binary variable for the cutoff
  mutate(fds_binary = case_when(fds_score >= 9 ~ "Yes",
                                fds_score <= 8 ~ "No"))%>%
  mutate(fds_binary = factor(fds_binary, levels = c("No", "Yes"), ordered = T))

# View a summary of the new variable
summary(analyticNSHAP$fds_score)
summary(analyticNSHAP$fds_binary)
```

Frequent anxiety symptoms is designated as score >= 8. Scoring was conducted following the provided guidelines. (Payne, 2014)
```{r Scoring for frequent anxiety symptoms}
analyticNSHAP <- analyticNSHAP %>%
  # Assign scores to each item
  mutate(felttens = case_when(felttens == 1 ~ 0, 
                              felttens == 2 ~ 1, 
                              felttens == 3 ~ 2,
                              felttens == 4 ~ 3))%>%
  mutate(fright = case_when(fright == 1 ~ 0, 
                              fright == 2 ~ 1, 
                              fright == 3 ~ 2,
                              fright == 4 ~ 3))%>%
  mutate(worry = case_when(worry == 1 ~ 0, 
                              worry == 2 ~ 1, 
                              worry == 3 ~ 2,
                              worry == 4 ~ 3))%>%
  mutate(relaxed = case_when(relaxed == 1 ~ 3,
                             relaxed == 2 ~ 2,
                             relaxed == 3 ~ 1,
                             relaxed == 4 ~ 0))%>%
  mutate(butterfly = case_when(butterfly == 1 ~ 0, 
                              butterfly == 2 ~ 1, 
                              butterfly == 3 ~ 2,
                              butterfly == 4 ~ 3))%>%
  mutate(restless = case_when(restless == 1 ~ 0, 
                              restless == 2 ~ 1, 
                              restless == 3 ~ 2,
                              restless == 4 ~ 3))%>%
  mutate(panic = case_when(panic == 1 ~ 0, 
                              panic == 2 ~ 1, 
                              panic == 3 ~ 2,
                              panic == 4 ~ 3))%>%
  # Create a score variable based on the sum
  mutate(fas_score = felttens + fright + worry + relaxed + butterfly + restless + panic)%>%
  # Create a binary variable based on the cutoff (>=8)
  mutate(fas_binary = case_when(fas_score >=8 ~ "Yes",
                                fas_score <=7 ~ "No"))%>%
  mutate(fas_binary = factor(fas_binary, levels = c("No", "Yes"), ordered = T))

# View a summary of the newly created variables
summary(analyticNSHAP$fas_score)
summary(analyticNSHAP$fas_binary)
```

Anxiety and depressive symptoms can co-occur and be more debilitating than either set of symptoms on its own. A composite mental health variable was created to capture this information, using the Frequent Depressive Symptoms and Frequent Anxiety Symptoms scales provided. This requires a complete case approach, as the FAS scale was administered in the leave-behind questionnaire and as such has a lower response rate. Those who did not respond to the FDS and/or FAS questionnaire will be removed from the study sample. 
```{r Composite Mental Health}
analyticNSHAP <- analyticNSHAP %>%
  # Filter out those with NA responses to FAS or FDS
  filter(!is.na(fas_binary) & !is.na(fds_binary))%>%
  mutate(compMenHealth = case_when(fas_binary == "Yes" & fds_binary == "No" ~ "Frequent Anxiety Symptoms",
                                    fas_binary == "No" & fds_binary == "Yes" ~ "Frequent Depressive Symptoms",
                                    fas_binary == "No" & fds_binary == "No" ~ "Neither Anxiety nor Depressive Symptoms",
                                    fas_binary == "Yes" & fds_binary == "Yes" ~ "Frequent Anxiety & Depressive Symptoms"))%>%
  mutate(compMenHealth = factor(compMenHealth, levels = c("Neither Anxiety nor Depressive Symptoms",
                                                       "Frequent Depressive Symptoms",
                                                       "Frequent Anxiety Symptoms",
                                                       "Frequent Anxiety & Depressive Symptoms"), 
                               ordered = T, 
                               labels =  c("Neither Anxiety nor Depressive Symptoms",
                                                       "Frequent Depressive Symptoms",
                                                       "Frequent Anxiety Symptoms",
                                                       "Frequent Anxiety & Depressive Symptoms")))

# View a summary of the variable
summary(analyticNSHAP$compMenHealth)
```
Survey guidelines propose a binary classification of marital status - unpartnered/unmarried and partnered/married (Waite 2021). The original variable is recoded accordingly.
```{r Dichotomize partnership}
analyticNSHAP <- analyticNSHAP %>%
  # Create a binary variable and set as an ordered factor
  mutate(MARITLST = case_when(
    MARITLST %in% "(1) married" ~ "partnered/married",
    MARITLST %in% "(2) living with a partner" ~ "partnered/married",
    MARITLST %in% "(3) separated" ~ "unpartnered/unmarried",
    MARITLST %in% "(4) divorced" ~ "unpartnered/unmarried",
    MARITLST %in% "(6) never married" ~ "unpartnered/unmarried",
    is.na(MARITLST) ~ as.character(MARITLST)  # Handle NA values and keep them unchanged
  )) %>%
  mutate(MARITLST = factor(MARITLST, levels = c("unpartnered/unmarried", "partnered/married"), ordered = T))

# summary
summary(analyticNSHAP$MARITLST)
```

AS complete case analysis is planned for the independent and dependent variables, an indicator variable is created to designate individuals who will be included. This will facilitate rudimentary missing data analysis.
```{r inclusion indicator}
analyticNSHAP <- analyticNSHAP %>%
  mutate(includeIndicator = ifelse(!is.na(compMenHealth) & !is.na(ADL_severity) & !is.na(IADL_severity), "Yes", "No"))%>%
  mutate(includeIndicator = factor(includeIndicator, levels = c("Yes", "No")))

summary(analyticNSHAP$includeIndicator)
```


The NHSAP is a weighted survey with the following parameters:
Individual weight - WEIGHT_ADJ
Strata = STRATUM
PSU = CLUSTER

A full survey design object will be created and then subset to the sample of interest for further analysis. The sample for this study must have a response to the composite mental health variable, ADL score and IADL score. 
```{r Create Survey Object}
# Create a survey design object for the entire data set
fullDesign <- svydesign(id    = ~CLUSTER,
                          strata  = ~STRATUM,
                          weights = ~WEIGHT_ADJ,
                          nest    = TRUE,
                          data    = analyticNSHAP)

# Create a subset object with only those who have data on mood, ADLs, and IADLS
analyticDesign <- subset(fullDesign, !is.na(compMenHealth), !is.na(ADL_severity), !is.na(IADL_severity))

# View a summary
summary(analyticDesign)
```
Count the number of cases used
```{r case count}
analyticNSHAP %>%
  filter(!is.na(compMenHealth), !is.na(ADL_severity), !is.na(IADL_severity))%>%
  count()

count(analyticNSHAP)
```
There are 1907 cases with data for all of the following variables - depressive symptoms, anxiety symptoms, ADL severity, and IADL severity.

Exploratory data analysis - test all continuous variables in the data for normality using the weighted survey object.
Continuous variables include age. While comorbidities (comorbid), ADL score, IADL score,  and compADL_score are also numeric, these are count data and variation will be best summarized with a mean and sd for descriptive purposes.
```{r EDA}
# Age histogram
gghistweight_svy(analyticDesign, AGE)+
  labs(title = "Age Distribution (Weighted)")

# Comorbidity count
gghistweight_svy(analyticDesign, comorbid)+
  labs(title = "Count of Comorbidities (Weighted)")

# ADL severity
gghistweight_svy(analyticDesign, ADL_severity)+
  labs(title = "ADL Severity (Weighted)")

# ADL severity by mental health
ggboxweight_svy(analyticDesign, ADL_severity)+
  labs(title = "ADL Severity (Weighted)")+
  facet_grid(rows = "compMenHealth")+
  theme_minimal()

# IADL severity
gghistweight_svy(analyticDesign, IADL_severity)+
  labs(title = "IADL Severity (Weighted)")

# IADL severity by mental health
ggboxweight_svy(analyticDesign, IADL_severity)+
  labs(title = "IADL Severity (Weighted)")+
  facet_grid(rows = "compMenHealth")+
  theme_minimal()

# ADL Score
gghistweight_svy(analyticDesign, ADL_score)+
  labs(title = "ADL Score (Weighted)")

# IADL Score
gghistweight_svy(analyticDesign, IADL_score)+
  labs(title = "IADL Score (Weighted)")

# CompADL score
gghistweight_svy(analyticDesign, compADL_score)+
  labs(title = "Summed ADL and IADL Score (Weighted)")
```
For the primary outcome - Neither ADL nor IADL severity demonstrates a normal distribution.

For the secondary outcome - As a large proportion of the ADL and IADL responses are 0 (i.e. individuals requiring help with neither ADLs nor IADLs) and the data is counts, a zero-inflated poisson regression model will be used. 

Table one - comparison of covariates stratified by mental health status (none, dep, anx, both)
The following variables were evaluate as non-normal and analyzed as such: age
```{r table one}
# Create a vector of table one variables
table1vars <- c("AGE", "GENDER", "ETHGRP", "EDUC", "MARITLST", "PHYSHLTH", "comorbid", "HEARN_RECODE", "OTHERINSUR", "LACKMEDCARE", "LACKPRESMED")

# Create table of demographic information by mental health status using only the percentages
DemographicsTableOne <- svyCreateTableOne(data = analyticDesign, 
                                       strata = "compMenHealth",
                                       vars = table1vars, 
                                       includeNA = F, 
                                       addOverall = T)

# View the results
TableOneExport <- print(DemographicsTableOne, 
                        showAllLevels = T, 
                        missing = T, 
                        format = "p",
                        varLabels = T,
                        nonnormal = "AGE")
```
```{r export table one}
write.csv(TableOneExport, "WrittenExam_Table1.csv")
```

Predictors are significant with the exception of age. There may be relationships between the predictors, as several measure similar information (for instance, self-assessed physical health and the count of comorbidities). If predictors are not independent, they likely present redundant information and one representative variable should be carried through adjusted models.
```{r EDA - correlations in predictors}
# significance of self-assessed physical health and comorbidities - Kruskal Wallis
# Box plot of comorbidities by each group
ggboxweight_svy(analyticDesign, comorbid)+
  facet_grid(rows = "PHYSHLTH")

svyranktest(comorbid ~ PHYSHLTH,
            analyticDesign,
            method = "KruskalWallis")


```
The KruskalWallis test was significance for the comparison of comorbidity count across self-rated physical health groups. Thus, only one of the two variables should be included. As number of comorbidities is a standard measure and more objectively determined, it will be included. Self-rated physical health is likely also related to anxiety and depressive symptoms, as the perception of health may be affected by presence of frequent anxiety and depressive symptoms.

Do not have same variables in round 2 for lackpresmed and lackmedcare
Table two - comparison of primary and secondary outcomes stratified by mental health status (none, dep, anx, both)
```{r table 2}
table2vars <- c("ADL_severity", "IADL_severity", "ADL_score", "IADL_score", "compADL_score", "compADL_ordinal")

# Create tableone object for table 2 outcomes
OutcomesTableTwo <- svyCreateTableOne(data = analyticDesign, 
                                       strata = "compMenHealth",
                                       vars = table2vars, 
                                       includeNA = F, 
                                       addOverall = T)

# View the results
TableTwoExport <- print(OutcomesTableTwo, 
                        showAllLevels = T, 
                        missing = T, 
                        format = "p",
                        varLabels = T)#,
                        #nonnormal = c("ADL_severity", "IADL_severity"))
```

```{r export table 2}
write.csv(TableTwoExport, "WrittenExam_TableTwo.csv")
```

Missing data analysis - a table is created to compare the individuals who have data for depression, anxiety, ADL, and IADL severity with those who lack data for any of these variables
```{r missing data analysis}
# Create a list of variable
MDAvars <- c("AGE", "GENDER", "ETHGRP", "EDUC", "MARITLST", "PHYSHLTH", "comorbid", "HEARN_RECODE", "OTHERINSUR", "LACKMEDCARE", "LACKPRESMED", "ADL_severity", "IADL_severity", "ADL_score", "IADL_score")

# Create a stratified table
MDATable <- svyCreateTableOne(data = fullDesign, 
                              strata = "includeIndicator",
                              vars = MDAvars,
                              includeNA = F)
                             # addOverall = T)

# View the results
MDATableExport <- print(MDATable, 
                        showAllLevels = F, 
                        missing = T, 
                        format = "p",
                        varLabels = T)

# Export the results
write.csv(MDATableExport, "WrittenExams_CrudeMissingData.csv")
```


For the primary outcome - separate weighted linear regresison models will be fit for severity of difficulty with ADLs and IADLs. Unadjusted models with the composite mental health variables as the sole predictor and adjusted models accounting for predictors will be fit and compared. Odds ratios will be calculated along with significance of the coefficients. 
Adjusted models will include significantly different covariates between mental health groups, accounting for results of testing for relationship between predictors done previously (i.e. both comorbidity count and self-rated physical health were significantly different but the two were demonstrated to be related by the KruskalWallis test, thus only comorbidity count is included). These variables include: AGE, GENDER, EDUC, MARITLST, comorbid, HEARN_RECODE, ANYPRIVATEINS
```{r linear regresison - severity of ADL difficulty}
# Unadjusted model
mod.ADL_severity <- svyglm(ADL_severity ~ compMenHealth, design = analyticDesign)
mod.ADL_severity

mod.ADL_severity_summary <- summary(mod.ADL_severity)

# Adjusted model
mod.ADL_severity_adj <- svyglm(ADL_severity ~ compMenHealth + AGE + GENDER + EDUC + MARITLST + 
                          comorbid + HEARN_RECODE + OTHERINSUR,
                          design = analyticDesign)

mod.ADL_severity_adj_summary <- summary(mod.ADL_severity_adj)

# Calculate 95% CI for OR - not appropriate, need beta estimates
#ADL_severityCI <- exp(confint(mod.ADL_severity))
#ADL_severity_adjCI <- exp(confint(mod.ADL_severity_adj))

# Calculate odds ratios
# Point estimates
#ADL_severityOR <- exp(coef(mod.ADL_severity))
#ADL_severity_adjOR <- exp(coef(mod.ADL_severity_adj))

```

```{r linear regression - severity of IADL difficulty}
# Unadjusted model
mod.IADL_severity <- svyglm(IADL_severity ~ compMenHealth, design = analyticDesign)
mod.IADL_severity

mod.IADL_severity_summary <- summary(mod.IADL_severity)

# Adjusted model
mod.IADL_severity_adj <- svyglm(IADL_severity ~ compMenHealth + AGE + GENDER + EDUC + MARITLST + 
                          comorbid + HEARN_RECODE + OTHERINSUR,
                          design = analyticDesign)

mod.IADL_severity_adj_summary <- summary(mod.IADL_severity_adj)

# Calculate odds ratios - not appropriate, need beta estimates
#IADL_severityOR <- exp(coef(mod.IADL_severity))
#IADL_severity_adjOR <- exp(coef(mod.IADL_severity_adj))

# Calculate 95% CI
#IADL_severityCI <- exp(confint(mod.IADL_severity))
#IADL_severity_adjCI <- exp(confint(mod.IADL_severity_adj))
```

```{r export linear regression results}
# Unadjusted model
ADLseverity_unadjustedTbl <- data.frame(ADLBeta = mod.ADL_severity_summary$coefficients[,1], 
                                        ADL_StdError = mod.ADL_severity_summary$coefficients[,2],
                                        ADL_t = mod.ADL_severity_summary$coefficients[,3],
                                        ADLp = mod.ADL_severity_summary$coefficients[,4], 
                                        IADLBeta = mod.IADL_severity_summary$coefficients[,1],
                                        IADL_StdError = mod.IADL_severity_summary$coefficients[,2],
                                        IADL_t = mod.IADL_severity_summary$coefficients[,3],
                                        IADLp = mod.IADL_severity_summary$coefficients[,4])

# Adjusted Model
ADLseverity_AdjustedTbl <- data.frame(ADLBeta = mod.ADL_severity_adj_summary$coefficients[,1], 
                                      ADL_StdError = mod.ADL_severity_adj_summary$coefficients[,2],
                                        ADL_t = mod.ADL_severity_adj_summary$coefficients[,3],
                                        ADLp = mod.ADL_severity_adj_summary$coefficients[,4], 
                                        IADLBeta = mod.IADL_severity_adj_summary$coefficients[,1], 
                                      IADL_StdError = mod.IADL_severity_adj_summary$coefficients[,2],
                                        IADL_t = mod.IADL_severity_adj_summary$coefficients[,3],
                                        IADLp = mod.IADL_severity_adj_summary$coefficients[,4])

# Export tables
write.csv(ADLseverity_unadjustedTbl, "WrittenExams_UnadjustedLinRegADLSeverity.csv")
write.csv(ADLseverity_AdjustedTbl, "WrittenExams_AdjustedADLSeverity.csv")
```

As the AIC for the adjusted models is significantly lower than the unadjusted models, only the adjusted models will be used going forward.

Check assumptions for linear regression:
1. Observations are independent - assumed met
2. The Outcome is continuous - met (interval scale from 0-21)
3. Linearity - relationship between outcome and each continuous predictor is linear
      This should be tested for Age and number of comorbidities
4. Homoscedasticity - 
5. Independence of residuals
6. Normality of residuals
7. No perfect multicollinearity - this can be assessed with VIF, using the cutoff of GVIF^(1/2*df))<2.5
```{r lin reg assumption}

# Check for multicollinearity
car::vif(mod.ADL_severity_adj)
car::vif(mod.IADL_severity_adj)
```

Calculate model significance for linear regression models
could look at r-squared, unclear whether this is warranted with survey data
```{r lin reg significance}
#trying Jtools summ function
jtools::summ(mod.ADL_severity_adj, confint = TRUE, digits = 3)
#jtools::export_summs(mod.ADL_severity_adj)
#jtools::plot_summs(mod.ADL_severity)

#jtools::summ(mod.ADL_severity)

jtools::summ(mod.IADL_severity_adj, confint = TRUE, digits = 3)
```
Visualization of the multivariate linear regression model to use in poster, where beta estimates and 95% confidence intervals are plotted.
```{r linear regression visualization}
# create a vector of labels
plotLabels <- c("Covered by Private Insurance", "Income: $100k or Greater", "Income: $50,000 - $99,999", "Income: $25,000 - $49,999", "Number of Comorbidities",   "Partnered/Married", "Bachelor's Degree or Higher", "Some College", "High School Degree or Equivalent","Female", "Age", "Frequent Anxiety & Depressive Symptoms",  "Frequent Anxiety Symptoms", "Frequent Depressive Symptoms")

# Create a plot of the beta estimates using plot_summs from jtools
LinRegPlot <- plot_summs(mod.ADL_severity_adj, mod.IADL_severity_adj,
                   model.names = c("ADL Severity", "IADL Severity"), 
                   colors = "CUD Bright",
                   point.size = 12)+
                theme(text = element_text(size = 20, family = "Arial", colour = "black"),
                      axis.text.y = element_text(size = 20, family = "Arial", colour = "black"))+
                scale_y_discrete(labels = plotLabels)
# Save the plot
ggsave("LinRegForestPlot.png", LinRegPlot, width = 18, height = 10)
```